Uncertainty in Inflation Forecasting Using Bayesian Methods

### 2022-05-10 draft_nb.Rmd
5/10 update 1

---
title: "Uncertainty in Inflation Forecasting Using Bayesian Methods"
author: Samuel Louissaint
output: html_document
---
# Introduction
The project will be using some macroeconomic knowledge, specifically the Monetarist understanding of economic activity. It's defined by a deceivingly simply formula: ${MV = PQ}$  
According to this definition, M is the money supply, V is the money velocity, P is the price of goods and services, and Q is the quantity of goods and services. Meaning the price of goods can be demonstrated as such: ${P = \frac{MV}{Q}}$  
In terms of economic indicators measured by the Federal Reserve Bank, this formula is best represented by:  
**P - Consumer Price Index for All Urban Consumers: All Items in U.S. City Average (though for this investigation we'll be using the percentage change from previous year figures to represent inflationary movements)**  
**M - Real M2 Money Stock**  
**V - Velocity of M2 Money Stock**  
**Q - Real Gross Domestic Product**  
Luckily, the Fed saved me the trouble of adjusting these figures for CPI myself, and for that I thank them deeply. In terms of a scientific model, there isn't an independent variable we can manipulate freely like in a controlled experiment. However, there is a tool regularly used by the Fed to temper inflation and unemployment, the interest rate. According to macroeconomic theory, increasing interest rates decreases lending/borrowing and thus consumer spending, decreasing inflation at the cost of economic growth. Because it is the only indicator that is so closely related to inflation ***and*** can be directly manipulated --- though often as a response to other economic activity or fiscal policy changes --- it serves as the exposure variable in the following directed acyclic graph (DAG).
```{r, echo=FALSE, results="hide", warning=FALSE}
packages <- c('ggdag', 'dagitty', 'tidyverse', 'tidybayes', 'bsts', 'rstan', 'modelr', 'brms', 'gganimate', 'cowplot', 'ggridges', 'colorspace', 'lubridate', 'rethinking')
suppressPackageStartupMessages(lapply(packages, library, character.only = T))
```

>>>>

---
title: "Uncertainty in Inflation Forecasting Using Bayesian Methods"
author: Samuel Louissaint
output:
  github_document:
    toc: true
    toc_depth: 2
    fig_width: 5
    fig_height: 5
---
# Introduction
The project will be using some macroeconomic knowledge, specifically the Monetarist understanding of economic activity. It's defined by a deceivingly simply formula: ${MV = PQ}$  
According to this definition, M is the money supply, V is the money velocity, P is the price of goods and services, and Q is the quantity of goods and services. Meaning the price of goods can be demonstrated as such: ${P = \frac{MV}{Q}}$  
In terms of economic indicators measured by the Federal Reserve Bank, this formula is best represented by:  
**P - Consumer Price Index for All Urban Consumers: All Items in U.S. City Average (though for this investigation we'll be using the percentage change from previous year figures to represent inflationary movements)**  
**M - Real M2 Money Stock**  
**V - Velocity of M2 Money Stock**  
**Q - Real Gross Domestic Product**  
Luckily, the Fed saved me the trouble of adjusting these figures for CPI myself, and for that I thank them deeply. In terms of a scientific model, there isn't an independent variable we can manipulate freely like in a controlled experiment. However, there is a tool regularly used by the Fed to temper inflation and unemployment, the interest rate. According to macroeconomic theory, increasing interest rates decreases lending/borrowing and thus consumer spending, decreasing inflation at the cost of economic growth. Because it is the only indicator that is so closely related to inflation ***and*** can be directly manipulated --- though often as a response to other economic activity or fiscal policy changes --- it serves as the exposure variable in the following directed acyclic graph (DAG).
```{r, echo=FALSE, results="hide", warning=FALSE}
packages <- c('ggdag', 'dagitty', 'tidyverse', 'tidybayes', 'bsts', 'rstan', 'modelr', 'brms', 'gganimate', 'cowplot', 'ggridges', 'colorspace', 'lubridate')
new.packages <- packages[!(packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
suppressPackageStartupMessages(lapply(packages, library, character.only = T))
```
